### 2024-11-28 paper/paper.qmd
Update Measurement

---
title: "Assessing Bridge Conditions in New York State: The Impact of Age, Municipality, and Structural Deficiencies"
subtitle: "My subtitle if needed"
author: 
  - Ariel Xing
thanks: "Code and data are available at: [https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis](https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true 
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

The dataset, titled New_York_Bridges_2016[@2016data], was orignaly posted by[@NYS], collected by [@2016data]. Original website now are with newly updated data in 2020. This data contains detailed information about bridges in New York State. The raw dataset has 17502 rows and 13 columns, capturing structural, geographical, and condition-related aspects of 17502 bridges. After cleaning the data, we keep 4 variables for analysis, whith response variable "Condition" and three predictor variables, "AgeAtInspection"(numeric), "Located_Municipality"(categorical) and "Owner_Goup"(categorical).

A similar dataset, New York Bridges 2020[@2020data], was considered but not used because its key variable for condition is binary (indicating whether the bridge is in poor condition or not). While this binary variable might be useful for certain analyses, it does not provide the granularity needed to fully understand bridge conditions or to assess the relationships between explanatory variables and condition scores. The 2016 dataset, by contrast, includes a numeric measure of condition, allowing for a more nuanced analysis.

## Measurement

In New York State, bridge inspectors from the State Department of Transportation (NYSDOT) conduct evaluations of non-toll bridges every two years to ensure their safety and structural integrity. Their assessments are performed on a span-by-span basis, meaning that each section of the bridge between its supporting points is inspected individually. For every bridge, inspectors examine up to 47 structural elements, including 25 span-specific components such as the deck, beams, and supports. Each component is assigned a condition score, ranging from 1 (severely deficient) to 7 (new or excellent condition), which reflects its structural health at the time of inspection. Inspectors also note any defects, such as cracks, corrosion, or wear, and document associated quantities, such as the length, area, or number of affected elements, to capture the extent of the damage.

In addition to these detailed evaluations, inspectors assign federal ratings based on the overall average condition of a bridge’s major components. These federal ratings correspond to the values of the **Condition** variable used in our paper and analysis. Bridges with condition scores below 5 are classified as being in “poor” condition and are further categorized as either Structurally Deficient (SD) or Functionally Obsolete (FO). All findings from these assessments are recorded in inspection reports, providing essential data to guide maintenance, rehabilitation, or replacement efforts.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References

### 2024-11-28 scripts/03-clean_data.R
delete redundant package

@@ -9,7 +9,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(stringr)
 library(arrow)
 
 #### Clean data ####

### 2024-11-28 scripts/03-clean_data.R
Make AgeAtInspection in cleaned data as numeric variable

@@ -22,6 +22,23 @@ raw_data <- read_delim(
   guess_max = 20000                                         # Improves guessing of column types by sampling more rows
 )
 
+# Ensure the necessary columns are numeric and drop rows with missing values
+raw_data <- raw_data %>%
+  mutate(
+    YearInspected = as.numeric(YearInspected),    # Convert YearInspected to numeric for calculations
+    AgeAtInspection = as.numeric(AgeAtInspection), # Convert AgeAtInspection to numeric for calculations
+    Built = as.numeric(Built) ) %>%                    # Convert Built to numeric for comparison
+  drop_na()         
+
+# Filter the data to check for bridges that may have been rebuilt.
+# This identifies cases where the calculated construction year (YearInspected - AgeAtInspection)
+# does not match the recorded "Built" year, potentially indicating a rebuild or data inconsistency.
+filtered_data <- raw_data %>%
+  filter(round(YearInspected - AgeAtInspection) != Built)
+
+# Count how many rows were identified as potential rebuilds or inconsistencies.
+nrow(filtered_data)
+
 # Normalize Municipality names
 raw_data <- raw_data %>%
   mutate(LastWord = str_extract(Municipality, "\\w+$")) %>%  # Add a column with the last word of Municipality

### 2024-11-28 scripts/03-clean_data.R
Remove useless code

@@ -22,22 +22,10 @@ raw_data <- read_delim(
   guess_max = 20000                                         # Improves guessing of column types by sampling more rows
 )
 
-# Ensure the necessary columns are numeric and drop rows with missing values
+# Ensure the necessary column is numeric and drop rows with missing values
 raw_data <- raw_data %>%
-  mutate(
-    YearInspected = as.numeric(YearInspected),    # Convert YearInspected to numeric for calculations
-    AgeAtInspection = as.numeric(AgeAtInspection), # Convert AgeAtInspection to numeric for calculations
-    Built = as.numeric(Built) ) %>%                    # Convert Built to numeric for comparison
+  mutate(AgeAtInspection = as.numeric(AgeAtInspection)) %>% # Convert AgeAtInspection to numeric for calculations
   drop_na()         
-
-# Filter the data to check for bridges that may have been rebuilt.
-# This identifies cases where the calculated construction year (YearInspected - AgeAtInspection)
-# does not match the recorded "Built" year, potentially indicating a rebuild or data inconsistency.
-filtered_data <- raw_data %>%
-  filter(round(YearInspected - AgeAtInspection) != Built)
-
-# Count how many rows were identified as potential rebuilds or inconsistencies.
-nrow(filtered_data)
 
 # Normalize Municipality names
 raw_data <- raw_data %>%

### 2024-11-28 scripts/06-model_data.R
Delete useless code

@@ -18,18 +18,11 @@ library(arrow)
 bridge_data <- read_parquet("data/02-analysis_data/cleaned_bridge_condition_data.parquet")
 
 ### Model data ####
-# Ensure categorical variables are treated as factors for the model
-bridge_data <- bridge_data %>%
-  mutate(Located_Municipality = as.factor(Located_Municipality),
-         Owner_Group = as.factor(Owner_Group))
-
-# Ensure AgeAtInspection is treated as a numeric variable
-bridge_data <- bridge_data %>%
-  mutate(AgeAtInspection = as.numeric(AgeAtInspection))
-
 
 # Fit a linear model with AgeAtInspection , Located_Municipality and Owner_Group
 model <- lm(Condition ~ AgeAtInspection + Located_Municipality + Owner_Group, data = bridge_data)
+
+summary(model)
 
 #### Save model ####
 # Save the model object for future use

### 2024-11-28 paper/paper.qmd
Upload data section

@@
 library(tidyverse)
-library(palmerpenguins)
+library(ggplot2)
+library(arrow)
+library(knitr)
+bridge_data <- read_parquet("../data/02-analysis_data/cleaned_bridge_condition_data.parquet")
 ```
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The dataset used in this analysis, titled “New_York_Bridges_2016”, was accessed from the Data And Story Library (DASL) [@dasl_data_2016], which archives historical data originally compiled by the New York State Open Data platform [@ny_open_data_2020].
+The original dataset includes contributions from county-level data provided by the New York State Department of Transportation (NYSDOT) [@county_data].
 
-The dataset, titled New_York_Bridges_2016[@2016data], was orignaly posted by[@NYS], collected by [@2016data].
-Original website now are with newly updated data in 2020.
-This data contains detailed information about bridges in New York State.
-The raw dataset has 17502 rows and 13 columns, capturing structural, geographical, and condition-related aspects of 17502 bridges.
-After cleaning the data, we keep 4 variables for analysis, whith response variable "Condition" and three predictor variables, "AgeAtInspection"(numeric), "Located_Municipality"(categorical) and "Owner_Goup"(categorical).
+This dataset was chosen over the updated 2020 version from New York State Open Data due to the change in how bridge condition data is recorded.
+In the 2016 dataset, the condition variable is a continuous variable measured on a seven-point scale (1 = "Deficient Condition" to 7 = "Good Condition").
+However, in the 2020 dataset, NYSDOT replaced this scale with a binary Poor/OK variable, recorded as Poor01 (1 = Poor, 0 = OK).
+This change limits the analytical potential of the dataset, as the binary variable lacks the granularity required for linear regression that relies on continuous variables to detect nuanced trends or incremental differences.
+By using the 2016 dataset, this study retains the detailed numerical condition ratings, enabling to analysis impact of factors such as age, location, and owners on bridge conditions.
 
-A similar dataset, New York Bridges 2020[@2020data], was considered but not used because its key variable for condition is binary (indicating whether the bridge is in poor condition or not).
-While this binary variable might be useful for certain analyses, it does not provide the granularity needed to fully understand bridge conditions or to assess the relationships between explanatory variables and condition scores.
-The 2016 dataset, by contrast, includes a numeric measure of condition, allowing for a more nuanced analysis.
+The raw dataset contains 17,502 rows and 13 columns, providing structural, geographical, and condition-related data on bridges across New York State.
+After data cleaning, four key variables were selected for analysis: the response variable “Condition” and three predictor variables, “AgeAtInspection”, “Located_Municipality”, and “Owner_Group”.
+The variables “Condition” and “AgeAtInspection” are numeric and were directly retained from the raw data.
+The categorical variables “Located_Municipality” and “Owner_Group” were newly constructed by categorizing the original variables “Municipality” and “Owner”, respectively.
+
+To conduct the analysis for this paper, we utilized the statistical programming language R [@citeR].
+The workflow was supported by several packages, including tidyverse [@tidyverse] for data cleaning and manipulation, arrow [@arrow] for efficient data storage and retrieval, validate [@validate] for data testing and validation, knitr[@knitr] for creating tables and ggplot2 [@ggplot2] for creating visualizations.
 
@@
 Bridges with condition scores below 5 are classified as being in “poor” condition and are further categorized as either Structurally Deficient (SD) or Functionally Obsolete (FO).
-All findings from these assessments are recorded in inspection reports, providing essential data to guide maintenance, rehabilitation, or replacement efforts.
+All findings from these assessments are recorded in inspection reports, providing essential data to guide maintenance, rehabilitation, or replacement efforts.[@measurement]
 
-## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Outcome Variable
+### Condition 
+The **Condition** variable reflects the overall state of a bridge, encompassing both its structural integrity and functional adequacy as determined during inspections.
+It is measured on a continuous scale from 1 to 7, with higher scores indicating better condition.
+Bridges with condition scores above 5 are considered to be in good standing, while scores below 5 indicate that a bridge may be classified as either Structurally Deficient (SD) or Functionally Obsolete (FO).
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-condition-summary
+#| tbl-cap: "Summary Statistics of NYS Bridge Condition Ratings"
 #| echo: false
+#| tbl-pos: "H"
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Summary statistics for Condition
+condition_summary <- bridge_data %>%
+  summarise(
+    Min = min(Condition, na.rm = TRUE),
+    Max = max(Condition, na.rm = TRUE),
+    Mean = mean(Condition, na.rm = TRUE),
+    Median = median(Condition, na.rm = TRUE),
+    "Standard Deviation" = sd(Condition, na.rm = TRUE),
+    Count = n()
+  )
+
+# Create a table
+kable(condition_summary, format = "latex", booktabs = TRUE)
 ```
+@tbl-condition-summary presents for the condition ratings of bridges in New York State, based on a dataset of 15,855 observations.
+The mean condition rating of 5.38 indicates that, on average, bridges are in fair to good condition.
+The median rating of 5.30 suggests that half of the bridges have condition ratings below or equal to 5.30.
+The close alignment of the mean and median reflects a relatively symmetrical distribution.
+While the minimum condition rating is 1.93, and the maximum is 7, the standard deviation of 0.83 suggests that most bridges cluster around the mean, with only a few outliers at the extreme ends of the scale.
+These statistics indicate that the majority of bridges are maintained at a stable and adequate level, with relatively few in very poor or excellent condition.
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+\newpage
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-condition-histogram
+#| fig-cap: "Distribution of Bridge Condition Ratings"
 #| echo: false
@@
 #| message: false
+#| Fig-pos: "H"
+
+# Create a histogram for the Condition variable
+ggplot(bridge_data, aes(x = Condition)) +
+  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
+  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(1, 7)) + # Ensures x-axis shows 1 to 7
+  labs(
+    x = "Condition",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
+
+@fig-condition-histogram The histogram displays the distribution of bridge condition ratings in New York State.
+The majority of bridges are clustered around the middle to upper end of the condition scale, with relatively few bridges at either the lower or higher ends.
+The left tail, representing bridges in severely poor condition, is thin, indicating that such bridges are uncommon.
+Similarly, the right tail, corresponding to bridges in excellent condition, is sparse, suggesting that high-quality, newly constructed bridges are not predominant.
+The distribution highlights a concentration of bridges in moderate condition, with a gradual decrease in frequency toward both ends, reflecting the natural aging process and variability in maintenance practices across the state.
+
+## Predictor variables
+### AgeAtInspection
+The **AgeAtInspection** variable is a continuous measure that represents the age of a bridge in years.
+It is calculated as the time elapsed from the year the bridge was originally built or entirely replaced to its most recent inspection date.
+
+\newpage
+```{r}
+#| label: fig-age-histogram
+#| fig-cap: "Distribution of Bridge Ages at the Time of Inspection. The red dashed lines mark intervals of 25 years and the blue density curve provides a smooth visualization of the distribution"
+#| echo: false
+#| Fig-pos: "H"
+#| warning: false
+#| message: false
+
+# Visualization: Histogram with mode markers and density curve
+ggplot(bridge_data, aes(x = AgeAtInspection)) +
+  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) + # Histogram
+  geom_vline(xintercept = seq(25, 150, by = 25), color = "red", linetype = "dashed", size = 0.8) + # Vertical lines for modes
+  geom_density(aes(y = ..count..), color = "blue", size = 1) + # Density curve scaled to histogram
+  labs(
+    x = "Age at Inspection (Years)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
 
 ```
+@fig-age-histogram shows distribution of bridge ages at the time of inspection, overlaid with red dashed lines and a blue density curve.
+The red dashed lines mark intervals of 25 years and align with clear peaks in the histogram, suggesting periodic patterns in bridge construction or major renovations.
+Prominent clusters are observed around 25, 50, and 75 years of age, indicating that many bridges were likely built or reconstructed during these timeframes.
+The blue density curve provides a smooth visualization of the distribution, confirming the periodic modes while highlighting the overall trends.
+Beyond 100 years, the frequency drops significantly, with very few bridges older than this still in operation, reflecting the typical lifecycle of infrastructure.
+The concentration of bridges between 20 and 80 years old suggests that a significant portion of bridges are mid-life and may soon require maintenance or replacement.@fig-age-histogram highlights both the historical patterns of bridge construction and the current age distribution of bridges in the dataset.
+```{r}
+#| label: fig-ageatinspection-condition
+#| fig-cap: "Relationship Between AgeAtInspection and Condition.
+Each blue dot represents a bridge, showing its age at the time of inspection and the corresponding condition rating obtained from that inspection."
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"
 
-Talk way more about it.
+# Visualization: Relationship between AgeAtInspection and Condition
+ggplot(bridge_data, aes(x = AgeAtInspection, y = Condition)) +
+  geom_point(alpha = 0.3, color = "blue") +
+  labs(
+    x = "Age at Inspection (Years)",
+    y = "Condition"
+  ) +
+  theme_minimal()
+```
+@fig-ageatinspection-condition illustrates the relationship between bridge age at inspection and their condition ratings.
+Most points are concentrated in the upper-left quadrant, showing that younger bridges tend to have higher condition ratings (close to 7).
+As bridge age increases, condition ratings generally decline, reflecting the expected deterioration over time.
+However, beyond approximately 80 to 100 years, the relationship becomes less defined, with condition ratings displaying a wide range.
+This variability may be influenced by differences in maintenance practices, varying levels of transportation load, or disparities in construction quality among older bridges.
+Overall, the pattern indicates that bridge age is a key factor in determining condition, particularly for younger and midlife bridges.
 
-## Predictor variables
+### Located_Municipality
+The **Located_Municipality** variable categorizes bridges based on the type of municipality where they are located. The three categories are City, Town, and Village.
 
-Add graphs, tables and text.
+```{r}
+#| label: tbl-located-municipality-proportion
+#| tbl-cap: "Proportion of Bridges by Municipality.
+The first column shows the type of municipality (Located_Municipality), and the second column represents the proportion of bridges in each category (Percentage Account)."
+#| echo: false
+#| tbl-pos: "H"
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+# Proportion of bridges by municipality
+located_municipality_proportion <- bridge_data %>%
+  group_by(Located_Municipality) %>%
+  summarise(
+    "Percentage Account" = paste0(round((n() / nrow(bridge_data)) * 100, 2), "%")
+  )
+
+# Create a formatted table
+kable(located_municipality_proportion, format = "latex", booktabs = TRUE)
+```
+@tbl-located-municipality-proportion shows the distribution of bridges across different municipality types.
+Towns account for the majority of the bridges, representing 82.11% of the dataset, while cities and villages make up much smaller proportions, at 9.15% and 8.75%, respectively.
+This indicates that most bridges are located in towns, likely due to their larger geographic coverage.
+```{r}
+#| label: tbl-located-municipality-condition
+#| tbl-cap: "Average Bridge Condition by Municipality.
+The first column shows the type of municipality (Located_Municipality), and the second column provides the average condition rating (Average Condition) for bridges in each category."
+#| echo: false
+#| tbl-pos: "H"
+
+# Average condition by municipality
+located_municipality_condition <- bridge_data %>%
+  group_by(Located_Municipality) %>%
+  summarise(
+    "Average Condition" = round(mean(Condition, na.rm = TRUE), 2)
+  )
+
+# Create a formatted table
+kable(located_municipality_condition, format = "latex", booktabs = TRUE)
+```
+@tbl-located-municipality-condition presents the average condition ratings of bridges by municipality type.
+Bridges in towns have the highest average condition rating of 5.41, followed by villages at 5.33 and cities at 5.16.
+The variation in average condition ratings may reflect differences in maintenance practices, traffic loads, or resource allocation among the municipality types.
+
+### Owner_Group
+The **Owner_Group** variable categorizes bridges based on their ownership.
+This variable is designed to estimate potential differences in bridge conditions that may arise from varying ownership responsibilities.
+The categories include Municipalities, NYSDOT, and Other.
+The Other category combines bridges owned by private entities, authorities, commissions, and similar groups, as each accounts for a small portion of the data, making it practical to group them together
+```{r}
+#| label: tbl-owner-group-proportion
+#| tbl-cap: "Proportion of Bridges by Ownership.
+The first column shows the type of ownership group (Owner_Group), and the second column represents the proportion of bridges in each category (Percentage Account)."
+#| echo: false
+#| tbl-pos: "H"
+
+# Proportion of bridges by ownership
+owner_group_proportion <- bridge_data %>%
+  group_by(Owner_Group) %>%
+  summarise(
+    "Percentage Account" = paste0(round((n() / nrow(bridge_data)) * 100, 2), "%")
+  )
+
+# Create a formatted table
+kable(owner_group_proportion, format = "latex", booktabs = TRUE)
+```
+@tbl-owner-group-proportion summarizes the ownership distribution of bridges.
+The two largest ownership groups are municipalities and NYSDOT, which account for 49.83% and 43.05% of the dataset, respectively.
+The remaining 7.13% of bridges fall under the "Other" category, which includes private, authority, and commission ownerships.
+The high percentage of bridges owned by municipalities and NYSDOT reflects their dominant role in managing public infrastructure.
+
+```{r}
+#| label: tbl-owner-group-condition
+#| tbl-cap: "Average Bridge Condition by Ownership.
+The first column shows the type of ownership group (Owner_Group), and the second column provides the average condition rating (Average Condition) for bridges in each category."
+#| echo: false
+#| tbl-pos: "H"
+
+# Average condition by ownership
+owner_group_condition <- bridge_data %>%
+  group_by(Owner_Group) %>%
+  summarise(
+    "Average Condition" = round(mean(Condition, na.rm = TRUE), 2)
+  )
+
+# Create a formatted table
+kable(owner_group_condition, format = "latex", booktabs = TRUE)
+```
+@tbl-owner-group-condition shows the average condition ratings of bridges grouped by ownership.
+Bridges owned by municipalities have the highest average condition rating of 5.45, while NYSDOT-owned bridges have a slightly lower rating of 5.33.
+Bridges in the "Other" ownership category have the lowest average condition rating of 5.10, which may reflect less consistent maintenance or differing priorities among private or non-government entities.
 

### 2024-11-29 paper/paper.qmd
Add revision of Data section and upload draft of Model Section.

@@
 It is measured on a continuous scale from 1 to 7, with higher scores indicating better condition.
-Bridges with condition scores above 5 are considered to be in good standing, while scores below 5 indicate that a bridge may be classified as either Structurally Deficient (SD) or Functionally Obsolete (FO).
+Bridges with condition scores above 5 are considered to be in good standing, while scores below 5 indicate that a bridge is classified as either Structurally Deficient (SD) or Functionally Obsolete (FO).
 
@@
   summarise(
-    Min = min(Condition, na.rm = TRUE),
-    Max = max(Condition, na.rm = TRUE),
-    Mean = mean(Condition, na.rm = TRUE),
-    Median = median(Condition, na.rm = TRUE),
-    "Standard Deviation" = sd(Condition, na.rm = TRUE),
+    Min = round(min(Condition, na.rm = TRUE), 2),
+    Max = round(max(Condition, na.rm = TRUE), 2),
+    Mean = round(mean(Condition, na.rm = TRUE), 2),
+    Median = round(median(Condition, na.rm = TRUE), 2),
+    "Standard Deviation" = round(sd(Condition, na.rm = TRUE), 2),
     Count = n()
@@
 ```
-@tbl-condition-summary presents for the condition ratings of bridges in New York State, based on a dataset of 15,855 observations.
+@tbl-condition-summary presents for the condition ratings of bridges in New York State, based on a clenaed dataset of 15,855 observations.
 The mean condition rating of 5.38 indicates that, on average, bridges are in fair to good condition.
@@
   labs(
-    x = "Condition",
+    x = "Bridge Condition Rating",
     y = "Frequency"
@@
 ## Predictor variables
-### AgeAtInspection
+### Age of Bridge at Inspection
 The **AgeAtInspection** variable is a continuous measure that represents the age of a bridge in years.
@@
   labs(
-    x = "Age at Inspection (Years)",
+    x = "Bridge Ages at Inspection (Years)",
     y = "Frequency"
@@
 Beyond 100 years, the frequency drops significantly, with very few bridges older than this still in operation, reflecting the typical lifecycle of infrastructure.
-The concentration of bridges between 20 and 80 years old suggests that a significant portion of bridges are mid-life and may soon require maintenance or replacement.@fig-age-histogram highlights both the historical patterns of bridge construction and the current age distribution of bridges in the dataset.
+The concentration of bridges between 20 and 80 years old suggests that a significant portion of bridges are mid-life and may soon require maintenance or replacement.It highlights both the historical patterns of bridge construction and the current age distribution of bridges in the dataset.
 ```{r}
 #| label: fig-ageatinspection-condition
-#| fig-cap: "Relationship Between AgeAtInspection and Condition.
+#| fig-cap: "Relationship Between Bridge Age at Inspection and Bridge Condition Rating.
 Each blue dot represents a bridge, showing its age at the time of inspection and the corresponding condition rating obtained from that inspection."
@@
   labs(
-    x = "Age at Inspection (Years)",
-    y = "Condition"
+    x = "Bridge Age at Inspection (Years)",
+    y = "Bridge Condition Rating"
   ) +
@@
 
-### Located_Municipality
+### Bridge Located Municipality
 The **Located_Municipality** variable categorizes bridges based on the type of municipality where they are located. The three categories are City, Town, and Village.
@@
 #| tbl-cap: "Proportion of Bridges by Municipality.
-The first column shows the type of municipality (Located_Municipality), and the second column represents the proportion of bridges in each category (Percentage Account)."
+The first column lists the type of municipality where the bridges are located, including City, Town, and Village, while the second column shows the proportion of bridges in each category."
 #| echo: false
@@
 
+# Rename the first column of the summarize table
+colnames(located_municipality_proportion)[
+  colnames(located_municipality_proportion) == "Located_Municipality"
+] <- "Bridge Located Municipality"
+
 # Create a formatted table
@@
 #| label: tbl-located-municipality-condition
-#| tbl-cap: "Average Bridge Condition by Municipality.
-The first column shows the type of municipality (Located_Municipality), and the second column provides the average condition rating (Average Condition) for bridges in each category."
+#| tbl-cap: "Average Bridge Condition Rating by Municipality.
+The first column lists the type of municipality where the bridges are located, including City, Town, and Village, and the second column provides the average condition rating for bridges in each category."
 #| echo: false
@@
   summarise(
-    "Average Condition" = round(mean(Condition, na.rm = TRUE), 2)
+    "Average Condition Rating" = round(mean(Condition, na.rm = TRUE), 2)
   )
+
+# Rename the first column of the summarize table
+colnames(located_municipality_condition)[
+  colnames(located_municipality_condition) == "Located_Municipality"
+] <- "Bridge Located Municipality"
 
@@
 
-### Owner_Group
+### Ownership of Bridge
 The **Owner_Group** variable categorizes bridges based on their ownership.
@@
 #| tbl-cap: "Proportion of Bridges by Ownership.
-The first column shows the type of ownership group (Owner_Group), and the second column represents the proportion of bridges in each category (Percentage Account)."
+The first column lists the type of bridge ownership, including NYSDOT, Municipality, and Other (private entities, authorities, commissions), while the second column shows the proportion of bridges in each category."
 #| echo: false
@@
   )
+
+# Rename the first column of the summarize table
+colnames(owner_group_proportion)[
+  colnames(owner_group_proportion) == "Owner_Group"
+] <- "Ownership of Bridge"
 
@@
 #| label: tbl-owner-group-condition
-#| tbl-cap: "Average Bridge Condition by Ownership.
-The first column shows the type of ownership group (Owner_Group), and the second column provides the average condition rating (Average Condition) for bridges in each category."
+#| tbl-cap: "Average Bridge Condition Rating by Ownership.
+The first column lists the type of bridge ownership, including NYSDOT, Municipality, and Other (private entities, authorities, commissions), and the second column provides the average condition rating for bridges in each category."
 #| echo: false
@@
   summarise(
-    "Average Condition" = round(mean(Condition, na.rm = TRUE), 2)
+    "Average Condition Rating" = round(mean(Condition, na.rm = TRUE), 2)
   )
+
+# Rename the first column of the summarize table
+colnames(owner_group_condition)[
+  colnames(owner_group_condition) == "Owner_Group"
+] <- "Ownership of Bridge"
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modeling strategy is to investigate the relationship between bridge condition ratings and key predictors, including **AgeAtInspection**, **Located_Municipality**, and **Owner_Group**.
+The model aims to identify significant patterns and evaluate how these factors influence bridge conditions.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here, we present a linear regression model, explaining its setup, assumptions, and justification. Details of model diagnostics and validation can be found in [Appendix -@sec-model-details].
 
-## Model set-up
+## Model Setup
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the condition rating of bridge $i$, measured on a scale from 1 to 7, where higher values represent better condition. The predictors are defined as follows:
+- $x_{1i}$: Age of bridge $i$ at the time of inspection (**AgeAtInspection**), measured in years.
+- $x_{2i}$: Type of municipality where the bridge is located (**Located_Municipality**), categorized as **City**, **Town**, or **Village**.
+- $x_{3i}$: Type of ownership of bridge $i$ (**Owner_Group**), categorized as **Municipality**, **NYSDOT**, or **Other**.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+The linear regression model is expressed as:
+```{=latex}
+\[
+\begin{aligned}
+y_i &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i, \text{City}} + \beta_3 x_{2i, \text{Village}} \\
+&\quad + \beta_4 x_{3i, \text{NYSDOT}} + \beta_5 x_{3i, \text{Other}} + \epsilon_i, \\
+\epsilon_i &\sim \text{Normal}(0, \sigma^2),
+\end{aligned}
+\]
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+The model assumes that the relationship between predictors and the response variable is linear, and that the residuals are normally distributed with constant variance.
 
-```
+### Software Implementation
+
+We implemented the model in R using the `lm()` function, which fits a linear regression model using ordinary least squares.
+Model validation and diagnostics, including residual analysis and goodness-of-fit metrics, were conducted to assess model performance.
+
+## Model Justification
+
+The inclusion of **AgeAtInspection** as a continuous predictor reflects its importance in capturing the natural deterioration of bridges over time.
+**Located_Municipality** and **Owner_Group** were included as categorical predictors to account for variations in maintenance practices, funding, and infrastructure quality that differ across municipalities and ownership types.
+These variables align directly with those defined in the data section, ensuring consistency between the data description and modeling approach.
+
+We chose a linear regression model for its simplicity, interpretability, and appropriateness for the continuous response variable (**Condition**).
+While other models, such as generalized additive models or Bayesian regressions, could capture non-linearities or incorporate priors, the current dataset supports a linear approach without overcomplicating the analysis.
+
+### Assumptions and Limitations
+
+The model assumes:
+- A linear relationship between predictors and the response variable.
+- Homoscedasticity (constant variance of residuals).
+- Normality of residuals.
+- Independence of observations.
+
+Potential limitations include the exclusion of interaction terms and other factors (e.g., traffic volume, environmental conditions) that could influence bridge condition.
+Furthermore, the linear regression model may not capture complex non-linear relationships or spatial dependencies inherent in the data.
+
+### Validation and Diagnostics
+
+To validate the model, we performed:
+- **Residual Analysis**: Checked for patterns in residuals to confirm assumptions of linearity and homoscedasticity.
+- **Goodness-of-Fit Metrics**: Calculated $R^2$ and adjusted $R^2$ to evaluate how well the model explains variability in the data.
+- **Sensitivity Analysis**: Assessed the impact of excluding extreme outliers on the estimated coefficients.
+- **Alternative Models**: Considered models with interaction terms and different transformations of predictors but found no substantial improvement in model performance.
+
+Details of model diagnostics, including plots of residuals and additional goodness-of-fit metrics, are provided in [Appendix -@sec-model-details].
+
+# Results
 

### 2024-12-01 paper/paper.qmd
Update for model part

@@
 library(arrow)
+library(broom)
 library(knitr)
@@
 To conduct the analysis for this paper, we utilized the statistical programming language R [@citeR].
-The workflow was supported by several packages, including tidyverse [@tidyverse] for data cleaning and manipulation, arrow [@arrow] for efficient data storage and retrieval, validate [@validate] for data testing and validation, knitr[@knitr] for creating tables and ggplot2 [@ggplot2] for creating visualizations.
+The workflow was supported by several packages, including tidyverse [@tidyverse] for data cleaning and manipulation, arrow [@arrow] for efficient data storage and retrieval, validate [@validate] for data testing and validation, knitr[@knitr] for creating tables and ggplot2 [@ggplot2] and broom[@broom] for creating visualizations.
 
@@
 
-The goal of our modeling strategy is to investigate the relationship between bridge condition ratings and key predictors, including **AgeAtInspection**, **Located_Municipality**, and **Owner_Group**.
-The model aims to identify significant patterns and evaluate how these factors influence bridge conditions.
+The goal of our modeling strategy is examine the relationship between **bridge condition ratings(condition)** and key predictors, including **the type of municipality where the bridge is located (Located_Municipality)**, **the age of the bridge at the time of inspection (AgeAtInspection)**, and **the ownership of the bridge (Owner_Group)**.
+By constructing a multiple linear regression model, we aim to quantify the impact of these factors on changes in bridge conditions.
+This analysis will provide a foundation for estimating the time and budget required for future bridge repair projects, supporting effective infrastructure maintenance planning.
 
-Here, we present a linear regression model, explaining its setup, assumptions, and justification. Details of model diagnostics and validation can be found in [Appendix -@sec-model-details].
+Here, we present a multiple linear regression model, explaining its setup, assumptions, and justification. Details of model diagnostics and validation can be found in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the condition rating of bridge $i$, measured on a scale from 1 to 7, where higher values represent better condition. The predictors are defined as follows:
-- $x_{1i}$: Age of bridge $i$ at the time of inspection (**AgeAtInspection**), measured in years.
-- $x_{2i}$: Type of municipality where the bridge is located (**Located_Municipality**), categorized as **City**, **Town**, or **Village**.
-- $x_{3i}$: Type of ownership of bridge $i$ (**Owner_Group**), categorized as **Municipality**, **NYSDOT**, or **Other**.
+The Multiple linear regression model, **Bridge Condition model**,  is expressed as:
 
-The linear regression model is expressed as:
-```{=latex}
-\[
-\begin{aligned}
-y_i &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i, \text{City}} + \beta_3 x_{2i, \text{Village}} \\
-&\quad + \beta_4 x_{3i, \text{NYSDOT}} + \beta_5 x_{3i, \text{Other}} + \epsilon_i, \\
-\epsilon_i &\sim \text{Normal}(0, \sigma^2),
-\end{aligned}
-\]
+$$
+\text{Condition}_i = \beta_0 + \beta_1 \cdot \text{Age}_i 
+    + \beta_2 \cdot I(\text{Municipality} = \text{City}) 
+    + \beta_3 \cdot I(\text{Municipality} = \text{Village})
+$$
+$$
+    + \beta_4 \cdot I(\text{Owner} = \text{NYSDOT}) 
+    + \beta_5 \cdot I(\text{Owner} = \text{Other}) 
+    + \epsilon_i
+$$
+Here:
 
-```
+- $\text{Condition}_i$: Condition rating of bridge $i$ (response variable).
+- $\text{Age}_i$: Age of bridge $i$ at the time of inspection.
+- $I(\text{Municipality} = \text{City})$: Indicator variable for bridges located in a **City**.
+- $I(\text{Municipality} = \text{Village})$: Indicator variable for bridges located in a **Village**.
+  - Reference category: **Town**.
+- $I(\text{Owner} = \text{NYSDOT})$: Indicator variable for bridges owned by **NYSDOT**.
+- $I(\text{Owner} = \text{Other})$: Indicator variable for bridges owned by **Other entities**.
+  - Reference category: **Municipalities**.
+- $\beta_0$: Intercept, representing the average condition rating for a Town-owned bridge with zero age.
+- $\beta_1$: Coefficient for the effect of Age on condition rating.
+- $\beta_2, \beta_3$: Coefficients for the effect of municipality type (City or Village, relative to Town).
+- $\beta_4, \beta_5$: Coefficients for the effect of ownership type (NYSDOT or Other, relative to Municipality).
+- $\epsilon_i$: Error term, capturing unexplained variation.
 
@@
 
-### Software Implementation
-
-We implemented the model in R using the `lm()` function, which fits a linear regression model using ordinary least squares.
-Model validation and diagnostics, including residual analysis and goodness-of-fit metrics, were conducted to assess model performance.
+We implemented the multiple linear regression model in R using the lm() function, which estimates coefficients through ordinary least squares.
 
 ## Model Justification
+### Variable Choice
+The selection of predictors in the model reflects key aspects discussed in the data section, ensuring a comprehensive understanding of bridge condition ratings.
+Age at Inspection was included as a continuous variable because @fig-ageatinspection-condition demonstrated a clear relationship between bridge age and condition ratings.
+Using age as a continuous predictor avoids the loss of information that would occur if age were categorized into groups, allowing the model to capture subtle variations in condition ratings as bridges age.
+Additionally, this choice aligns with the dataset's structure, where age is measured on a continuous scale.
 
-The inclusion of **AgeAtInspection** as a continuous predictor reflects its importance in capturing the natural deterioration of bridges over time.
-**Located_Municipality** and **Owner_Group** were included as categorical predictors to account for variations in maintenance practices, funding, and infrastructure quality that differ across municipalities and ownership types.
-These variables align directly with those defined in the data section, ensuring consistency between the data description and modeling approach.
+Municipality Type (City, Town, Village) was treated as a categorical variable with Town as the reference category.
+The data revealed that condition ratings vary systematically between municipality types, likely due to differences in traffic loads, maintenance resources, and environmental stressors.
+Modeling Municipality Type as levels rather than using dummy variables for individual municipalities ensures clarity and interpretability, highlighting differences between City, Town, and Village bridges.
 
-We chose a linear regression model for its simplicity, interpretability, and appropriateness for the continuous response variable (**Condition**).
-While other models, such as generalized additive models or Bayesian regressions, could capture non-linearities or incorporate priors, the current dataset supports a linear approach without overcomplicating the analysis.
+Ownership Type (Municipalities, NYSDOT, Other) was similarly treated as a categorical variable, allowing the model to reflect systematic differences in bridge conditions based on ownership responsibilities.
+The inclusion of ownership accounts for varying maintenance schedules and resource allocation.
+The reference category (Municipalities) was chosen because it comprises the largest proportion of bridges in the dataset, providing a stable baseline for comparison.
 
-### Assumptions and Limitations
+Together, these features were selected to balance the need for interpretability and explanatory power while reflecting meaningful patterns observed in the data section.
 
-The model assumes:
-- A linear relationship between predictors and the response variable.
-- Homoscedasticity (constant variance of residuals).
-- Normality of residuals.
-- Independence of observations.
+### Model Choice
+A multiple linear regression model was chosen for its simplicity, interpretability, and suitability for quantifying the influence of individual predictors while controlling for others.
+This model effectively supports the research goal of understanding how Age, Municipality Type, and Ownership Type affect bridge condition ratings.
 
-Potential limitations include the exclusion of interaction terms and other factors (e.g., traffic volume, environmental conditions) that could influence bridge condition.
-Furthermore, the linear regression model may not capture complex non-linear relationships or spatial dependencies inherent in the data.
+The model assumes a linear relationship between the predictors and the response variable (Condition Rating).
+This assumption is reasonable given observed trends, including the clear negative association between Age and Condition, and systematic differences between categories of Municipality and Ownership.
+While this assumption aligns with the data, it may not fully capture non-linear relationships or interactions between predictors.
+Nevertheless, the simplicity of linear regression is appropriate for this study’s goals, balancing clarity with analytical rigor.
 
-### Validation and Diagnostics
+Other key assumptions include Normality of Residuals and homoscedasticity (constant variance of residuals).
+Diagnostic tools were used to validate these assumptions, with details provided in [Appendix -@sec-model-diagnostics].
 
-To validate the model, we performed:
-- **Residual Analysis**: Checked for patterns in residuals to confirm assumptions of linearity and homoscedasticity.
-- **Goodness-of-Fit Metrics**: Calculated $R^2$ and adjusted $R^2$ to evaluate how well the model explains variability in the data.
-- **Sensitivity Analysis**: Assessed the impact of excluding extreme outliers on the estimated coefficients.
-- **Alternative Models**: Considered models with interaction terms and different transformations of predictors but found no substantial improvement in model performance.
+### Limitations
+Several limitations must be acknowledged.
+Due to data constraints, the model excludes potentially influential factors such as transport load, weather conditions, and environmental influences, introducing the risk of omitted variable bias.
+This limitation reduces the explanatory power of the model and may skew the estimated effects of included predictors.
+Additionally, the model does not account for interaction effects, such as how age and municipality might jointly influence condition ratings.
+Incorporating such interactions could enhance the model’s explanatory ability but would increase its complexity.
+Finally, the linearity assumption, while reasonable, may oversimplify more nuanced relationships in the data.
 
-Details of model diagnostics, including plots of residuals and additional goodness-of-fit metrics, are provided in [Appendix -@sec-model-details].
+Despite these limitations, the model's moderate complexity ensures it is neither overly simplistic nor unnecessarily complicated, making it appropriate for addressing the research question.
+
+## Alternative model
+In addition to the multiple linear regression model, we considered a generalized additive model (GAM), which unique strengths but were ultimately deemed less suitable for this analysis compared to multiple linear regression.
+
+A GAM could have been used to capture potential non-linear relationships between Age and Condition Rating.
+This flexibility is particularly beneficial when predictors exhibit non-linear effects, which might better reflect the natural deterioration of bridges over time.
+However, while GAMs provide excellent predictive performance, they sacrifice interpretability, as the model outputs are more challenging to communicate to non-technical audiences.
+Given the project's emphasis on explaining the relationship between predictors and condition ratings in a transparent manner, this complexity outweighed the benefits.
 
@@
 
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+# Model Diagnostics {#sec-model-diagnostics}
+## Linearity and Homoscedasticity Check
 
 ```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| label: fig-residual-plot
+#| fig-cap: "Residuals vs.
+Fitted Values for Bridge Condition rating Model.
+The residuals (differences between observed and predicted bridge condition ratings) are on the y-axis and the fitted values (predicted condition ratings from the model) are on the x-axis.
+Each dot represents a bridge, showing the magnitude and direction of its residual.
+The red dashed line indicates the zero residual line, representing where residuals would lie if the predictions were perfect."
+#| fig-pos: "H"
+#| echo: False
 
+# Fit the model
+model <- lm(Condition ~ AgeAtInspection + Located_Municipality + Owner_Group, data = bridge_data)
+
+# Residual Plot: Linearity and Homoscedasticity
+ggplot(data = augment(model), aes(.fitted, .resid)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  labs(x = "Fitted Values", y = "Residuals") +
+  theme_minimal()
 ```
+@fig-residual-plot provides evidence supporting the assumptions of linearity and homoscedasticity for the model.
+First, the random scatter of residuals around the horizontal line at zero indicates no systematic patterns, suggesting that the relationship between the predictors and the response variable is approximately linear.
+This random distribution implies that the model captures the linear relationships effectively without significant curvature or deviations.
 
-## Diagnostics
+Second, while the upper residuals show a declining trend due to the bounded nature of the response variable (maximum rating of 7), the lower residuals and the bulk of residuals across the range of fitted values exhibit a relatively constant variance.
+This consistency in spread suggests that the assumption of homoscedasticity (constant variance of residuals) holds for the majority of the data, making the deviations in the upper range reasonable and contextually acceptable.
+Together, these observations indicate that the model adequately meets the assumptions of linearity and homoscedasticity.
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+## Nomality of Residuals
 
 ```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
+#| label: fig-qq-plot
+#| echo: False
+#| fig-cap: "Q-Q Plot of Standardized Residuals for Bridge Condition Model.
+The x-axis represents the theoretical quantiles from a standard normal distribution, and the y-axis represents the standardized residuals of the Bridge Condition model.
+Each dot represents a bridge, showing how its residual compares to the expected normal distribution.
+The black dotted line represents the theoretical relationship the residuals would follow if they were perfectly normally distribute."
+#| fig-pos: "H"
+plot(model, which = 2)
 ```
+@fig-qq-plot indicates that the residuals of the model generally align with the theoretical quantiles of a normal distribution, as most points lie close to the reference line.
+While there is slight deviation in the extreme tails, this is not unusual for larger datasets and does not significantly impact the overall assumption of normality.
+The observed residual pattern supports the suitability of the model for the data, as minor deviations in the tails are unlikely to undermine the reliability of hypothesis testing or predictions.
+Therefore, the model can be considered appropriate for analysis.
 

### 2024-12-01 paper/paper.qmd
Update for the result section

@@
 ---
-title: "Assessing Bridge Conditions in New York State: The Impact of Age, Municipality, and Structural Deficiencies"
-subtitle: "My subtitle if needed"
+title: "Assessing Bridge Condition in New York State Using Multiple Linear Regression"
+subtitle: "Age, Municipality, and Ownership as Factors Influencing Bridge Condition Ratings "
 author: 
@@
 $$
-Here:
+where:
 
@@
 Given the project's emphasis on explaining the relationship between predictors and condition ratings in a transparent manner, this complexity outweighed the benefits.
+\newpage
 
 # Results
+The linear regression model predicting bridge condition ratings estimates how factors： the bridge age at inspection, the located municipality and the ownership influence the condition of bridges.The model result is shown in @tbl-model-result
+```{r}
+#| label: tbl-model-result
+#| tbl-cap: "Summary Table of Bridge Condition Model"
+#| echo: false
+#| tbl-pos: "H"
+
+model <- lm(Condition ~ AgeAtInspection + Located_Municipality + Owner_Group, data = bridge_data)
+
+# Extract coefficients
+coefficients_table <- tidy(model)
+
+# Rename terms to match model equation notation
+coefficients_table$term <- gsub("\\(Intercept\\)", "Intercept", coefficients_table$term)
+coefficients_table$term <- gsub("AgeAtInspection", "Age", coefficients_table$term)
+coefficients_table$term <- gsub("Located_MunicipalityTown", "Municipality = Town", coefficients_table$term)
+coefficients_table$term <- gsub("Located_MunicipalityVillage", "Municipality = Village", coefficients_table$term)
+coefficients_table$term <- gsub("Owner_GroupNYSDOT", "Owner = NYSDOT", coefficients_table$term)
+coefficients_table$term <- gsub("Owner_GroupOther", "Owner = Other", coefficients_table$term)
+
+# Select columns to display and round numbers
+selected_table <- coefficients_table %>%
+  select(term, estimate) %>%
+  mutate(across(where(is.numeric), ~ round(., 2)))
+
+# Rename columns for clarity
+colnames(selected_table) <- c("Variables", "Coefficients", "Std. Error")
+
+# Display the updated table
+kable(selected_table, align = "c", format = "markdown")
+```
+
+The intercept, estimated at 6.15, represents the baseline condition rating for a bridge located in a city, owned by the municipality, and with an age of zero years.
+This is a useful reference point against which the effects of age, municipality, and ownership are compared.
+
+The coefficient for Age is -0.02, indicating that for each additional year a bridge ages, its condition rating decreases by 0.02 points on average.
+This result reflects the natural deterioration of infrastructure over time, as older bridges are more likely to experience wear and tear.
+@fig-ageatinspection-condition presented in the Data section illustrates this trend, showing a clear negative relationship between bridge age and condition.
+This visual evidence aligns with the regression model, where the Age coefficient (-0.02) confirms this downward trend.
+Although the effect of age on condition appears small on an annual basis, the cumulative impact over several decades can significantly compromise a bridge's integrity, highlights the importance of regular inspections and maintenance programs to mitigate the impact of aging on bridge conditions.
+
+The location of the bridge, represented by the municipality variable, also plays a significant role in condition ratings.
+Bridges located in towns and villages have higher condition ratings compared to those in cities, as indicated by the positive coefficients for Municipality = Town (0.17) and Municipality = Village (0.19).
+These findings suggest that urban bridges are more prone to deterioration, possibly due to heavier traffic loads or more complex structural demands.
+In contrast, bridges in towns and villages may face fewer stressors, resulting in better overall condition ratings.
+
+Ownership of the bridge is another significant factor affecting its condition.
+The model uses Municipality-owned bridges as the reference category.
+The results show that bridges owned by Municipality have the highest condition rating,  NYSDOT has condition ratings that are 0.09 points lower, while those owned by Other entities have ratings 0.16 points lower, on average.
+These differences may reflect variations in maintenance practices, funding availability, or operational priorities across ownership groups.
+Bridges owned by municipalities might benefit from closer proximity to local resources and management, whereas NYSDOT and other entities may face challenges such as broader jurisdictional responsibilities or limited funding for specific assets.
+
+Taken together, these results highlight how a combination of aging, geographic context, and administrative oversight influences bridge condition ratings.
+Bridges in cities or under the management of entities other than municipalities tend to have lower ratings, while those in towns, villages, or managed by municipalities are generally better maintained.
+This underscores the importance of tailoring maintenance strategies to address the specific challenges faced by different ownership groups and geographic locations.
+Future studies could enhance this analysis by exploring interactions between municipality and ownership or incorporating additional variables such as traffic volume, funding levels, or structural design characteristics to provide a deeper understanding of the factors driving bridge condition outcomes.
 
@@
 
-Second, while the upper residuals show a declining trend due to the bounded nature of the response variable (maximum rating of 7), the lower residuals and the bulk of residuals across the range of fitted values exhibit a relatively constant variance.
-This consistency in spread suggests that the assumption of homoscedasticity (constant variance of residuals) holds for the majority of the data, making the deviations in the upper range reasonable and contextually acceptable.
-Together, these observations indicate that the model adequately meets the assumptions of linearity and homoscedasticity.
+Second, while the upper residuals show a declining trend due to the bounded nature of the response variable (maximum rating of 7), the lower residuals and the majority of residuals across the range of fitted values exhibit relatively constant variance.
+For bridges with lower condition ratings (fitted values below approximately 4), residuals display more variability and are predominantly negative, indicating that the model overpredicts for poorly maintained bridges.
+Despite these deviations, the overall spread of residuals remains consistent for the majority of the data, suggesting that the assumption of homoscedasticity is largely satisfied.
+Therefore, the model is deemed appropriate for analyzing the dataset, though its limitations in predicting low-condition bridges should be noted.
 

### 2024-12-03 other/datasheet/datasheet.qmd
Upload datasheet

---
title: "Datasheet for 'New York Bridge 2016'"
author: 
  - Ariel Xing
thanks: "Code and data are available at: https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis."
date: today
date-format: long
format: pdf
number-sections: true
bibliography: "references.bib"

---
This datasheet was created using the R programming language [@citeR].

**Motivation**

The "New York Bridges 2016" dataset was created to document and assess the structural conditions and administrative details of bridges across New York State, supporting infrastructure management, maintenance prioritization, and policymaking. Developed by the New York State Department of Transportation (NYSDOT) and made publicly available through the Data and Story Library (DASL), this dataset addresses the need for standardized, comprehensive data on bridge deterioration influenced by geographic, administrative, and traffic-related factors. Funded as part of New York State’s infrastructure monitoring programs, the dataset serves as a vital resource for researchers, policymakers, and engineers, enabling data-driven decision-making to enhance public safety and infrastructure sustainability.

**Composition**

The New York Bridges 2016 dataset comprises records of 17,502 individual bridges across the state, each representing a single instance. These data points include structural condition ratings, administrative ownership details, inspection dates, geographic locations, and additional features such as construction year and status indicators like "Structurally Deficient" (SD) or "Functionally Obsolete" (FO). These attributes make the dataset ideal for evaluating bridge conditions, deterioration trends, and the effectiveness of maintenance strategies.

Each bridge record is detailed and standardized, but not all bridges have complete information. For example, certain records lack exact inspection dates or detailed ownership classifications. This incompleteness reflects the challenges in collecting data consistently across a large and diverse infrastructure network. The dataset includes attributes like Municipality, Owner, Age at Inspection, and Condition Rating, making it suitable for stratified analyses based on geographic and administrative categories.

The dataset is comprehensive rather than a sample, aiming to capture all operational bridges in New York State as of 2016. This ensures extensive geographic and administrative coverage, from urban centers with high traffic density to rural bridges with lower traffic loads. While the data is representative, rural bridges may have less frequent or less detailed inspections, potentially introducing a minor bias.

The dataset's primary target variable is the Condition Rating, a numeric measure ranging from 0 to 7 that assesses the structural integrity of each bridge. Relationships between instances, such as proximity or ownership clustering, are not explicitly encoded but can be inferred using fields like Municipality or Owner.

This dataset does not rely on external resources or links and does not contain sensitive or personally identifiable information. It is publicly available and curated for research and policy planning purposes, ensuring ethical and practical usability for stakeholders. With its rich details and broad scope, the dataset supports in-depth analysis of infrastructure trends across New York State.

**Collection process**

The dataset titled "New_York_Bridges_2016" was originally compiled by the New York State Open Data platform[@ny_open_data_2020], with county-level contributions from the New York State Department of Transportation [@county_data]. Accessed via the Data And Story Library [@dasl_data_2016], which archives historical datasets for academic and public use, it represents a comprehensive resource for understanding bridge conditions across New York State.

Data for each bridge instance was acquired through direct inspections and administrative records. Observable characteristics such as structural condition ratings, age at inspection, and administrative ownership were collected via standardized field inspections conducted by certified engineers and inspectors, adhering to federal and state protocols. Validation procedures included regular audits and cross-checks to ensure data accuracy and consistency.

The primary mechanisms for data collection included on-site manual inspections supported by tools like measurement instruments and inspection checklists. Administrative details, such as ownership, were sourced from official state databases. NYSDOT ensured the reliability of these processes through rigorous training programs for inspectors and periodic calibration of measurement tools. The dataset is comprehensive, capturing all operational bridges in the state as of 2016, rather than representing a sample, and thus no specific sampling strategy was required.

The data collection process involved certified civil engineers, state employees, and field inspectors employed by NYSDOT. These professionals adhered to standard government pay structures, with no involvement of external contractors or crowdworkers. The data was collected over the span of 2016, aligning the collection timeframe with the creation of the dataset, ensuring the information reflects real-world conditions at the time.

Ethical review processes were not explicitly documented, as the dataset pertains exclusively to public infrastructure and does not involve sensitive or personal information. Data was not collected directly from individuals or third-party sources, eliminating concerns regarding notification, consent, or revocation mechanisms. Additionally, since the dataset focuses solely on publicly accessible infrastructure, no privacy or ethical issues were identified that would necessitate a formal review.

In summary, the New York Bridges 2016 dataset was compiled through standardized field inspections and administrative records, validated by trained professionals, and ethically sound due to its exclusive focus on public infrastructure. Its methodology ensures high-quality, accurate data for research, policy planning, and infrastructure management, while DASL plays a critical role in archiving and maintaining this resource for public and academic use.

**Uses**

The **New York Bridges 2016 dataset** should not be used for tasks requiring real-time data analysis or dynamic decision-making, as it represents a static snapshot of bridge conditions as of 2016. It is unsuitable for assessing current structural risks or predicting immediate failures without incorporating updated data. Additionally, the dataset should not be used for commercial purposes that could conflict with public interests, such as privatization analyses aimed at reducing public access to infrastructure. Finally, it is not appropriate for applications requiring sensitive or individualized data, as it does not contain personal information or real-time monitoring details.

**Distribution**
The New York Bridges 2016 dataset is publicly available and distributed through the Data and Story Library (DASL) website. It is accessible as a downloadable file in common formats such as text or CSV, ensuring ease of use for researchers and policymakers. The dataset does not have a dedicated digital object identifier (DOI), but it is maintained under publicly available terms, which can be accessed directly from the DASL platform.

The dataset is distributed without any copyright or intellectual property restrictions. It falls under public domain policies, given its origins in government data provided by the New York State Department of Transportation. There are no associated fees or licensing requirements, making it freely accessible for non-commercial, academic, and policy-related use.

No third parties have imposed intellectual property-based restrictions or limitations on the dataset, and there are no export controls or regulatory restrictions associated with its use. As a dataset focused solely on public infrastructure, it is designed for open access to encourage widespread research and application in infrastructure planning and analysis.

**Maintenance**

The New York Bridges 2016 dataset was originally published by Open Data NY and later archived by the Data and Story Library (DASL) for academic and historical purposes. While the dataset is static and will not be updated on DASL, newer versions of the data are periodically published as entirely new datasets on the Open Data NY platform.

For questions regarding the dataset hosted on DASL, users can reach out by filling out the contact form available [here](https://dasl.datadescription.com/datafile/new-york-bridges-2016/?_sf_s=Bridge&_sfm_cases=4+59943). This ensures that inquiries about the archived dataset can be addressed promptly. However, for the latest data or updated versions, users are encouraged to visit the Open Data NY platform, which continues to provide up-to-date datasets on bridge conditions in New York State.

This dual hosting approach ensures that both historical and current datasets remain accessible for research, policy-making, and educational purposes, catering to a broad spectrum of user needs.

\newpage

# References

### 2024-12-03 other/datasheet/datasheet.qmd
Upload datasheet

@@
 ---
-This datasheet was created using the R programming language [@citeR].
+This datasheet is created using programming language [@citeR]
 
@@
 
-The "New York Bridges 2016" dataset was created to document and assess the structural conditions and administrative details of bridges across New York State, supporting infrastructure management, maintenance prioritization, and policymaking.
+The "New York Bridges 2016" dataset was created to document and assess the structural conditions and administrative details of bridges across New York State, supporting infrastructure management, maintenance prioritization, and policy-making.
 Developed by the New York State Department of Transportation (NYSDOT) and made publicly available through the Data and Story Library (DASL), this dataset addresses the need for standardized, comprehensive data on bridge deterioration influenced by geographic, administrative, and traffic-related factors.

### 2024-12-03 paper/paper.qmd
Update abstract, introduction part. Upload a draft of appendix

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study examines the factors influencing bridge conditions in New York State, focusing on the roles of age, municipality location, and ownership, using condition ratings provided by NYSDOT as a measure of bridge quality.
+Applying a multiple linear regression model, the analysis reveals that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the progressive impact of aging.
+On average, bridges located in towns and villages have condition ratings that are 0.17 and 0.19 points higher, respectively, than those in cities.
+Additionally, bridges owned by NYSDOT and other entities have condition ratings that are 0.09 and 0.16 points lower, respectively, than those owned by municipalities.
+These findings highlight the importance of targeted maintenance strategies that account for geographic and administrative factors to enhance infrastructure sustainability and safety.
+"
 format: pdf
@@
 
-Overview paragraph
+Bridges are fundamental to New York State’s infrastructure, supporting transportation, economic activity, and community connections.
+They span diverse regions, from bustling urban centers to remote rural areas, accommodating varied traffic loads and environmental conditions.
+Maintaining these structures is essential to ensure public safety, economic efficiency, and reliable mobility.
+However, challenges such as aging infrastructure, resource constraints, and differing administrative responsibilities highlight the importance of understanding the factors that affect bridge conditions.
 
-Estimand paragraph
+The estimand in this study is the effect of three key predictors: bridge age, municipality type, and ownership, on bridge condition ratings.
+Using a multiple linear regression model, we analyze condition ratings from 17,502 bridges in New York State to quantify how these factors influence bridge deterioration.
 
-Results paragraph
+The results indicate that age is a major driver of bridge deterioration, with condition ratings decreasing by 0.02 points for each additional year.
+On average, bridges in towns and villages have better condition ratings than those in cities, with scores that are 0.17 and 0.19 points higher, respectively.
+Ownership also affects condition ratings: municipally owned bridges perform best, while NYSDOT-owned bridges score 0.09 points lower, and bridges owned by other entities score 0.16 points lower.
+These findings demonstrate the combined influence of geographic location and administrative responsibility on bridge conditions.
 
-Why it matters paragraph
+This research highlights the importance of targeted maintenance strategies that address the specific needs of aging bridges across different geographic and administrative contexts.
+By identifying the primary factors affecting bridge conditions, the study supports more effective resource distribution and prioritization in infrastructure management.
+These results contribute to the long-term sustainability and safety of the state’s transportation network.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows.
+@sec-data provides an overview of the dataset and variables used in the analysis.
+@sec-model outlines the regression model, its assumptions, and limitations.
+@sec-result presents the results, while @sec-discussion discusses their implications and offers recommendations for future research.
+@sec-methodology introduces an idealized methodology for collecting bridge traffic flow data, and @sec-model-diagnostics details the diagnostics performed to validate the model.
 
@@
     x = "Bridge Condition Rating",
-    y = "Frequency"
+    y = "Count"
   ) +
@@
     x = "Bridge Ages at Inspection (Years)",
-    y = "Frequency"
+    y = "Count"
   ) +
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-result}
 The linear regression model predicting bridge condition ratings estimates how factors： the bridge age at inspection, the located municipality and the ownership influence the condition of bridges.The model result is shown in @tbl-model-result
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Additional data details
+# Appendix: Idealized Methodology for Collecting Bridge Traffic Flow Data
+
+## Overview
+The objective of this methodology is to design a robust system for collecting traffic flow data across New York State bridges.
+This data serves as a foundation for understanding the interaction between traffic intensity and bridge conditions.
+By employing rigorous sampling strategies and advanced collection techniques, this methodology ensures that the collected data is representative and actionable for policy and maintenance decisions.
+
+## Target Population
+The target population encompasses all operational bridges in New York State.
+These include bridges in urban, suburban, and rural areas, varying by traffic volume and ownership.
+Bridges were selected to account for:
+
+- **Geographic Coverage**: Urban centers with high traffic density, suburban transit hubs, and rural connectors.
+- **Administrative Ownership**: Municipal, NYSDOT, and privately owned bridges.
+- **Traffic Intensity**: Low-traffic rural bridges, moderate suburban thoroughfares, and high-traffic urban crossings.
+
+This definition ensures inclusivity, enabling the methodology to capture regional differences in traffic flow and their effect on bridge conditions.
+
+## Sampling Approach
+
+### Stratification Variables
+To ensure the data represents the diversity of bridges, stratified random sampling is applied. Stratification variables include:
+
+- **Municipality Type**: City, town, village.
+- **Ownership**: Municipal, state, private.
+- **Traffic Volume**: High, medium, low daily use.
+
+Each bridge within a stratum is given an equal chance of selection, ensuring proportional representation across categories.
+
+### Sample Size
+Using the formula for sample size determination:
+
+$$
+n = \frac{Z^2 \cdot p \cdot (1-p)}{e^2}
+$$
+
+Where:
+
+- \( Z = 1.96 \) for 95% confidence,
+- \( p = 0.5 \), assuming maximum variability,
+- \( e = 0.05 \), margin of error.
+
+A minimum of 384 bridges is required to ensure statistically reliable data for the entire state.
+
+## Data Collection Techniques
+
+### Automatic Vehicle Counters (AVC)
+- Deployed on selected bridges to continuously record vehicle counts and classifications.
+- Cost-effective for high-traffic bridges requiring sustained monitoring.
+
+### Camera-Based Systems
+- Capture detailed traffic patterns, including speed and vehicle types, using AI-enhanced analysis.
+- Effective for complex traffic conditions but requires privacy safeguards.
+
+### GPS Data Integration
+- Utilize anonymized GPS data to complement on-site measurements, especially for remote or less-monitored bridges.
+
+### Manual Surveys
+- Conducted periodically to validate automated data and address gaps, particularly in low-traffic or rural settings.
+
+## Data Validation and Quality Control
+
+### Attention Checks
+Automated systems are periodically calibrated to ensure accuracy.
+
+### Cross-Validation
+Data from multiple sources (e.g., AVC, GPS) is compared to identify discrepancies.
+
+### Consistency Checks
+Traffic patterns are reviewed for anomalies that may indicate errors or data corruption.
+
+## Challenges and Limitations
+
+### Coverage Bias
+Rural bridges may be underrepresented due to accessibility challenges. Tailored solutions are necessary to address this.
+
+### Ethical Concerns
+Privacy issues related to camera and GPS monitoring must be mitigated through anonymization protocols.
+
+### Budget Constraints
+High-cost technologies like AI-enabled cameras may limit widespread deployment.
+
+## Relevance to Study
+Incorporating traffic flow data enhances the predictive model of bridge conditions by linking usage intensity to structural deterioration.
+This data allows policymakers to prioritize maintenance based on a combination of bridge condition and traffic stress.
 

### 2024-12-03 other/datasheet/datasheet.qmd
Update contact link

@@
 
-For questions regarding the dataset hosted on DASL, users can reach out by filling out the contact form available [here](https://dasl.datadescription.com/datafile/new-york-bridges-2016/?_sf_s=Bridge&_sfm_cases=4+59943).
+For questions regarding the dataset hosted on DASL, users can reach out by filling out the contact form available [here](https://datadescription.com/).
 This ensures that inquiries about the archived dataset can be addressed promptly.

### 2024-12-04 paper/paper.qmd
Completed paper

@@
 
-## First discussion point {#sec-first-point}
+## Age Distribution Patterns {#sec-age-distribution}
+The distribution of bridge ages at inspection demonstrates distinct peaks approximately every 25 years, reflecting historical trends in bridge construction or renovation efforts.
+These periodic patterns likely correspond to infrastructure development initiatives or funding cycles at specific points in time.
+Understanding these trends is crucial for proactive resource allocation, as large clusters of aging bridges may require simultaneous maintenance or replacement.
+Planning for these future needs will help mitigate risks associated with concentrated structural deterioration.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Repair Prioritization and Budget Allocation
+The regression model highlights the critical factors of bridge age, municipality type, and ownership in determining condition ratings.
+Urban bridges and those owned by NYSDOT or private entities consistently exhibit lower condition ratings, likely due to heavier traffic loads and diverse maintenance practices.
+These findings underscore the need for targeted repair strategies.
+Municipalities, which generally achieve better condition ratings, could serve as a benchmark for best practices in resource utilization.
+Repair prioritization should integrate both structural condition and traffic intensity to maximize the effectiveness of budget allocations, ensuring that high-use and socioeconomically vital bridges receive timely attention.
 
-## Second discussion point
+## Model Limitations and Future Improvements
+The model's low \( R^2 \) value, 0.396, suggests it does not fully capture the range of factors influencing bridge conditions.
+Key omitted variables, such as traffic volume, environmental stressors, and material quality, may significantly impact bridge deterioration.
+Future models should incorporate these variables, which can be obtained from traffic monitoring systems, meteorological data, and engineering databases, to enhance predictive accuracy.
+Furthermore, the assumption of constant variance is invalid for bridges with a condition rating below 4.5, indicating the current model is unsuitable for severely damaged bridges.
+While these bridges represent a small proportion of the dataset, their exclusion limits the model's scope.
+Future research should develop a separate model for low-condition bridges to address their unique deterioration dynamics.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Challenges with Low-Condition Bridges
+Based on the residual plot, the model tends to underestimate the actual condition ratings for low-condition bridges, as indicated by the positive residuals.
+This underestimation may be due to the complexity of these bridges' circumstances.
+First, low-condition bridges may have undergone emergency repairs, weight restrictions, or other temporary measures to maintain functionality.
+These interventions are not reflected in the model, leading to an underestimation of their condition.
+Second, such bridges might experience compounded effects from environmental stressors, high traffic loads, or material degradation, which the current linear framework fails to account for.
 
-## Third discussion point
+Furthermore, measures such as weight restrictions might reduce the load on these bridges, artificially slowing further deterioration and creating a mismatch between their actual state and the model's predictions.
+This non-linear degradation pattern suggests that the current model has limitations in fitting low-condition bridges.
+Future studies could improve upon this by incorporating variables such as traffic restrictions, emergency maintenance records, and environmental stress factors.
+Additionally, employing non-linear models or hierarchical approaches tailored to these bridges may better capture their unique deterioration dynamics.
+These refinements could lead to more precise and effective maintenance strategies for low-condition bridges.
 
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+## Geographic and Administrative Disparities in Bridge Conditions
+The analysis reveals significant disparities in bridge conditions based on geographic location and administrative ownership.
+Bridges in urban areas tend to have lower condition ratings compared to those in towns and villages, likely due to higher traffic volumes, environmental stress, and more complex structural demands in urban settings.
+Similarly, bridges owned by NYSDOT or private entities perform worse than those managed by municipalities, suggesting differences in maintenance priorities and resource allocation.
+These disparities point to the need for a more equitable distribution of infrastructure funds and tailored maintenance strategies that address the specific challenges faced by different regions and ownership groups.
+For example, urban bridges may benefit from more frequent inspections and targeted upgrades to withstand heavy usage, while state-owned and privately managed bridges could require enhanced funding mechanisms or stricter maintenance policies to ensure their long-term sustainability.
+Recognizing and addressing these geographic and administrative variations will be critical for maintaining a resilient and equitable bridge infrastructure network.
 
@@
 
-# Appendix: Idealized Methodology for Collecting Bridge Traffic Flow Data
+# Appendix: Idealized Methodology for Collecting Bridge Traffic Flow Data {#sec-methodology}
 
 ## Overview
-The objective of this methodology is to design a robust system for collecting traffic flow data across New York State bridges.
-This data serves as a foundation for understanding the interaction between traffic intensity and bridge conditions.
-By employing rigorous sampling strategies and advanced collection techniques, this methodology ensures that the collected data is representative and actionable for policy and maintenance decisions.
+The objective of this methodology is to design a comprehensive system for collecting and analyzing traffic flow data across New York State bridges.
+This additional data addresses a critical gap in the current dataset, which does not include traffic intensity—a key factor influencing bridge deterioration.
+By integrating traffic flow data, the enhanced model aims to improve its explanatory power and provide actionable recommendations for bridge maintenance prioritization.
+Rigorous sampling strategies and advanced data collection techniques are employed to ensure that the data is both representative and reliable, supporting more nuanced policy and engineering decisions.
 
@@
 
-- **Geographic Coverage**: Urban centers with high traffic density, suburban transit hubs, and rural connectors.
-- **Administrative Ownership**: Municipal, NYSDOT, and privately owned bridges.
-- **Traffic Intensity**: Low-traffic rural bridges, moderate suburban thoroughfares, and high-traffic urban crossings.
+- **Geographic Coverage**: Urban areas with dense traffic flows, suburban regions with moderate traffic levels, and rural areas characterized by infrequent usage.
+This distribution ensures that both high-stress urban crossings and lesser-used rural bridges are accounted for.
 
-This definition ensures inclusivity, enabling the methodology to capture regional differences in traffic flow and their effect on bridge conditions.
+- **Administrative Ownership**: Bridges under municipal (local government), NYSDOT (state-level), and private ownership.
+This stratification accounts for variations in maintenance standards and resources available to different administrative entities.
+
+- **Traffic Intensity**: Bridges are analyzed based on their average daily traffic (ADT) volume as a continuous numeric variable.
+This approach allows for precise modeling of traffic intensity and its impact on bridge conditions, enabling the detection of subtle trends and relationships that may be overlooked with categorical groupings.
+By treating ADT as a numeric variable, the methodology provides a more granular and accurate representation of traffic flow patterns across bridges.
+
+By incorporating these dimensions, the target population captures the variability in bridge conditions and usage, enabling a comprehensive analysis of how traffic impacts structural integrity.
 
@@
 ### Stratification Variables
-To ensure the data represents the diversity of bridges, stratified random sampling is applied. Stratification variables include:
+To ensure that the sample accurately reflects the diversity of bridges in New York State, stratified random sampling is applied. The stratification variables are:
 
@@
 - **Ownership**: Municipal, state, private.
-- **Traffic Volume**: High, medium, low daily use.
+- **Traffic Volume**: Recoord average daily traffic volume as a continuous numeric variable
 
-Each bridge within a stratum is given an equal chance of selection, ensuring proportional representation across categories.
+This approach provides proportional representation across all strata, ensuring that data collection is neither skewed toward urban high-traffic areas nor neglects low-traffic rural bridges.
 
@@
 ### Automatic Vehicle Counters (AVC)
-- Deployed on selected bridges to continuously record vehicle counts and classifications.
-- Cost-effective for high-traffic bridges requiring sustained monitoring.
-
+- Description: Deployed on selected bridges to continuously record vehicle counts and classifications.
+- Advantages: Cost-effective for high-traffic bridges requiring sustained monitoring.
+- Limitations: Limited effectiveness for capturing complex traffic behaviors, such as turning patterns.
 ### Camera-Based Systems
-- Capture detailed traffic patterns, including speed and vehicle types, using AI-enhanced analysis.
-- Effective for complex traffic conditions but requires privacy safeguards.
-
+- Description: AI-enhanced cameras monitor traffic, capturing vehicle speed, type, and flow patterns. Data is processed to identify trends and anomalies.
+- Advantages: Detailed and adaptable to a wide range of traffic conditions.
+- Limitations: Higher cost and potential privacy concerns, requiring strict anonymization protocols.
 ### GPS Data Integration
-- Utilize anonymized GPS data to complement on-site measurements, especially for remote or less-monitored bridges.
+- Description: Anonymized GPS data from vehicle navigation systems and smartphones supplements on-site traffic data, particularly for remote bridges with low traffic.
+- Advantages: Wide coverage and minimal additional hardware costs.
+- Limitations: Relies on external data sources, which may introduce variability in data quality.
 
 ### Manual Surveys
-- Conducted periodically to validate automated data and address gaps, particularly in low-traffic or rural settings.
-
-## Data Validation and Quality Control
-
-### Attention Checks
-Automated systems are periodically calibrated to ensure accuracy.
+- Description: Periodic on-site observations validate automated systems and address gaps in data, especially for rural or low-traffic bridges.
+- Advantages: Provides ground-truth data for calibration.
+- Limitations: Resource-intensive and limited to periodic snapshots.
+## Data Validation and Quality Control.
 
 ### Cross-Validation
-Data from multiple sources (e.g., AVC, GPS) is compared to identify discrepancies.
+Traffic flow data collected from different sources, such as AVCs, cameras, and GPS, are compared to identify and resolve discrepancies.
 
 ### Consistency Checks
-Traffic patterns are reviewed for anomalies that may indicate errors or data corruption.
+Traffic patterns are reviewed for outliers and anomalies, such as sudden spikes or drops in volume, which may indicate data errors or temporary disruptions (e.g., road closures).
+
+### Calibration
+Manual survey results are used to calibrate automated systems, ensuring accuracy across different traffic conditions and bridge types.
 
@@
 ### Coverage Bias
-Rural bridges may be underrepresented due to accessibility challenges. Tailored solutions are necessary to address this.
+- Issue: Rural and remote bridges may be underrepresented due to logistical challenges and lower prioritization.
+- Mitigation: Oversampling and supplementary data collection methods, such as GPS integration, can address this imbalance.
 
 ### Ethical Concerns
-Privacy issues related to camera and GPS monitoring must be mitigated through anonymization protocols.
+- Issue: Privacy issues arise with camera-based and GPS monitoring systems.
+- Mitigation: Implement anonymization protocols and restrict data access to ensure compliance with ethical standards.
 
 ### Budget Constraints
-High-cost technologies like AI-enabled cameras may limit widespread deployment.
+- Issue: Advanced systems, such as AI-enabled cameras, require significant investment, potentially limiting their use.
+- Mitigation: Employ a hybrid approach combining cost-effective methods (AVCs, GPS) with targeted deployment of high-cost systems.
 

### 2024-12-04 other/datasheet/datasheet.qmd
Add a reference

@@
 
-The **New York Bridges 2016 dataset** should not be used for tasks requiring real-time data analysis or dynamic decision-making, as it represents a static snapshot of bridge conditions as of 2016.
+The New York State Bridge Conditions dataset, also utilized in the ARTBA's bridge report[@ARTBA], has been applied to analyze the structural conditions of bridges across New York.
+The dataset helps identify structurally deficient bridges, prioritize infrastructure projects, and allocate federal funds, such as the $2 billion bridge formula funds under the Infrastructure Investment and Jobs Act (IIJA).
+This dataset is well-suited for tasks like policy development, maintenance scheduling, risk assessment, and infrastructure planning, offering actionable insights into bridge management.
+However, its composition reflects current bridge conditions and may not fully account for recently completed repairs, which could influence its applicability for real-time analyses.
+The dataset is tailored to infrastructure evaluations and is not recommended for unrelated uses.
+Overall, it serves as a critical resource for understanding and improving bridge conditions in New York State.
+
+This dataset should not be used for tasks requiring real-time data analysis or dynamic decision-making, as it represents a static snapshot of bridge conditions as of 2016.
 It is unsuitable for assessing current structural risks or predicting immediate failures without incorporating updated data.
@@
 **Distribution**
+
 The New York Bridges 2016 dataset is publicly available and distributed through the Data and Story Library (DASL) website.

### 2024-12-04 paper/paper.qmd
Updated discussion

@@
 The blue density curve provides a smooth visualization of the distribution, confirming the periodic modes while highlighting the overall trends.
+This point will be further discussion in @sec-age-distribution.
 Beyond 100 years, the frequency drops significantly, with very few bridges older than this still in operation, reflecting the typical lifecycle of infrastructure.
@@
 ## Age Distribution Patterns {#sec-age-distribution}
-The distribution of bridge ages at inspection demonstrates distinct peaks approximately every 25 years, reflecting historical trends in bridge construction or renovation efforts.
-These periodic patterns likely correspond to infrastructure development initiatives or funding cycles at specific points in time.
-Understanding these trends is crucial for proactive resource allocation, as large clusters of aging bridges may require simultaneous maintenance or replacement.
-Planning for these future needs will help mitigate risks associated with concentrated structural deterioration.
+The distribution of bridge ages at inspection reveals distinct peaks approximately every 25 years, indicating a recurring cycle of bridge construction and major renovation.
+This pattern suggests that every 25 years, a significant number of bridges reach the point where extensive maintenance or replacement is necessary.
+The previous peak occurred roughly 25 years ago, meaning the next wave of large-scale repairs or new construction is approaching.
+Recognizing this pattern highlights the importance of early planning and budgeting to address the anticipated surge in infrastructure needs effectively, ensuring resources are ready to manage the simultaneous demands of aging bridges.
 
@@
 ## Model Limitations and Future Improvements
-The model's low \( R^2 \) value, 0.396, suggests it does not fully capture the range of factors influencing bridge conditions.
+The model's low $R^2$ value, 0.396, suggests it does not fully capture the range of factors influencing bridge conditions.
 Key omitted variables, such as traffic volume, environmental stressors, and material quality, may significantly impact bridge deterioration.

### 2024-12-04 paper/paper.qmd
Improve prose

@@
 abstract: "This study examines the factors influencing bridge conditions in New York State, focusing on the roles of age, municipality location, and ownership, using condition ratings provided by NYSDOT as a measure of bridge quality.
-Applying a multiple linear regression model, the analysis reveals that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the progressive impact of aging.
+Applying a multiple linear regression model, the analysis shows that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the progressive impact of aging.
 On average, bridges located in towns and villages have condition ratings that are 0.17 and 0.19 points higher, respectively, than those in cities.
@@
 However, in the 2020 dataset, NYSDOT replaced this scale with a binary Poor/OK variable, recorded as Poor01 (1 = Poor, 0 = OK).
-This change limits the analytical potential of the dataset, as the binary variable lacks the granularity required for linear regression that relies on continuous variables to detect nuanced trends or incremental differences.
+This change limits the analytical potential of the dataset, as the binary variable lacks the granularity required for linear regression that relies on continuous variables to detect subtle trends or incremental differences.
 By using the 2016 dataset, this study retains the detailed numerical condition ratings, enabling to analysis impact of factors such as age, location, and owners on bridge conditions.
@@
 ### Condition 
-The **Condition** variable reflects the overall state of a bridge, encompassing both its structural integrity and functional adequacy as determined during inspections.
+The **Condition** variable reflects the overall state of a bridge, capturing both its structural integrity and functional adequacy as determined during inspections.
 It is measured on a continuous scale from 1 to 7, with higher scores indicating better condition.
@@
 ### Variable Choice
-The selection of predictors in the model reflects key aspects discussed in the data section, ensuring a comprehensive understanding of bridge condition ratings.
+The selection of predictors in the model reflects key aspects discussed in the data section, ensuring a thorough understanding of bridge condition ratings.
 Age at Inspection was included as a continuous variable because @fig-ageatinspection-condition demonstrated a clear relationship between bridge age and condition ratings.
@@
 Municipality Type (City, Town, Village) was treated as a categorical variable with Town as the reference category.
-The data revealed that condition ratings vary systematically between municipality types, likely due to differences in traffic loads, maintenance resources, and environmental stressors.
+The data shows that condition ratings vary systematically between municipality types, likely due to differences in traffic loads, maintenance resources, and environmental stressors.
 Modeling Municipality Type as levels rather than using dummy variables for individual municipalities ensures clarity and interpretability, highlighting differences between City, Town, and Village bridges.
@@
 Incorporating such interactions could enhance the model’s explanatory ability but would increase its complexity.
-Finally, the linearity assumption, while reasonable, may oversimplify more nuanced relationships in the data.
+Finally, the linearity assumption, while reasonable, may oversimplify more subtle relationships in the data.
 
@@
 ## Age Distribution Patterns {#sec-age-distribution}
-The distribution of bridge ages at inspection reveals distinct peaks approximately every 25 years, indicating a recurring cycle of bridge construction and major renovation.
+The distribution of bridge ages at inspection shows distinct peaks approximately every 25 years, indicating a recurring cycle of bridge construction and major renovation.
 This pattern suggests that every 25 years, a significant number of bridges reach the point where extensive maintenance or replacement is necessary.
@@
 ## Repair Prioritization and Budget Allocation
-The regression model highlights the critical factors of bridge age, municipality type, and ownership in determining condition ratings.
+The regression model highlights the key factors of bridge age, municipality type, and ownership in determining condition ratings.
 Urban bridges and those owned by NYSDOT or private entities consistently exhibit lower condition ratings, likely due to heavier traffic loads and diverse maintenance practices.
@@
 ## Geographic and Administrative Disparities in Bridge Conditions
-The analysis reveals significant disparities in bridge conditions based on geographic location and administrative ownership.
+The analysis shows significant disparities in bridge conditions based on geographic location and administrative ownership.
 Bridges in urban areas tend to have lower condition ratings compared to those in towns and villages, likely due to higher traffic volumes, environmental stress, and more complex structural demands in urban settings.
@@
 For example, urban bridges may benefit from more frequent inspections and targeted upgrades to withstand heavy usage, while state-owned and privately managed bridges could require enhanced funding mechanisms or stricter maintenance policies to ensure their long-term sustainability.
-Recognizing and addressing these geographic and administrative variations will be critical for maintaining a resilient and equitable bridge infrastructure network.
+Recognizing and addressing these geographic and administrative variations will be essential for maintaining a resilient and equitable bridge infrastructure network.
 
@@
 ## Overview
-The objective of this methodology is to design a comprehensive system for collecting and analyzing traffic flow data across New York State bridges.
-This additional data addresses a critical gap in the current dataset, which does not include traffic intensity—a key factor influencing bridge deterioration.
+The objective of this methodology is to design a thorough system for collecting and analyzing traffic flow data across New York State bridges.
+This additional data addresses an essential limitation in the current dataset, which does not include traffic intensity—a key factor influencing bridge deterioration.
 By integrating traffic flow data, the enhanced model aims to improve its explanatory power and provide actionable recommendations for bridge maintenance prioritization.
-Rigorous sampling strategies and advanced data collection techniques are employed to ensure that the data is both representative and reliable, supporting more nuanced policy and engineering decisions.
+Rigorous sampling strategies and sophisticated data collection techniques are employed to ensure that the data is both representative and reliable, supporting more detailed policy and engineering decisions.
 
 ## Target Population
-The target population encompasses all operational bridges in New York State.
+The target population includes all operational bridges in New York State.
 These include bridges in urban, suburban, and rural areas, varying by traffic volume and ownership.
@@
 
-By incorporating these dimensions, the target population captures the variability in bridge conditions and usage, enabling a comprehensive analysis of how traffic impacts structural integrity.
+By incorporating these dimensions, the target population captures the variability in bridge conditions and usage, enabling a detailed analysis of how traffic impacts structural integrity.
 
@@
 ### Manual Surveys
-- Description: Periodic on-site observations validate automated systems and address gaps in data, especially for rural or low-traffic bridges.
+- Description: Periodic on-site observations validate automated systems and address missing information in data, especially for rural or low-traffic bridges.
 - Advantages: Provides ground-truth data for calibration.
@@
 ### Budget Constraints
-- Issue: Advanced systems, such as AI-enabled cameras, require significant investment, potentially limiting their use.
+- Issue:  High-cost systems, such as AI-enabled cameras, require significant investment, potentially limiting their use.
 - Mitigation: Employ a hybrid approach combining cost-effective methods (AVCs, GPS) with targeted deployment of high-cost systems.

### 2024-12-04 paper/paper.qmd
Final update

@@
 abstract: "This study examines the factors influencing bridge conditions in New York State, focusing on the roles of age, municipality location, and ownership, using condition ratings provided by NYSDOT as a measure of bridge quality.
-Applying a multiple linear regression model, the analysis shows that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the progressive impact of aging.
+Using a multiple linear regression model, the analysis indicates that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the cumulative effects of aging.
 On average, bridges located in towns and villages have condition ratings that are 0.17 and 0.19 points higher, respectively, than those in cities.
@@
 
-This research highlights the importance of targeted maintenance strategies that address the specific needs of aging bridges across different geographic and administrative contexts.
-By identifying the primary factors affecting bridge conditions, the study supports more effective resource distribution and prioritization in infrastructure management.
+This research emphasizes the need for targeted maintenance strategies tailored to the specific requirements of aging bridges across diverse geographic and administrative contexts.By identifying the primary factors affecting bridge conditions, the study supports more effective resource distribution and prioritization in infrastructure management.
 These results contribute to the long-term sustainability and safety of the state’s transportation network.
@@
 This change limits the analytical potential of the dataset, as the binary variable lacks the granularity required for linear regression that relies on continuous variables to detect subtle trends or incremental differences.
-By using the 2016 dataset, this study retains the detailed numerical condition ratings, enabling to analysis impact of factors such as age, location, and owners on bridge conditions.
+By using the 2016 dataset, this study retains detailed numerical condition ratings, enabling an analysis of the impact of factors such as age, location, and ownership on bridge conditions.
 
@@
 
-To conduct the analysis for this paper, we utilized the statistical programming language R [@citeR].
+To conduct the analysis for this paper, I utilized the statistical programming language R [@citeR].
 The workflow was supported by several packages, including tidyverse [@tidyverse] for data cleaning and manipulation, arrow [@arrow] for efficient data storage and retrieval, validate [@validate] for data testing and validation, knitr[@knitr] for creating tables and ggplot2 [@ggplot2] and broom[@broom] for creating visualizations.
@@
 
-The model assumes that the relationship between predictors and the response variable is linear, and that the residuals are normally distributed with constant variance.
+The model assumes a linear relationship between predictors and the response variable, with residuals normally distributed and exhibiting constant variance.
 
@@
 This visual evidence aligns with the regression model, where the Age coefficient (-0.02) confirms this downward trend.
-Although the effect of age on condition appears small on an annual basis, the cumulative impact over several decades can significantly compromise a bridge's integrity, highlights the importance of regular inspections and maintenance programs to mitigate the impact of aging on bridge conditions.
+Although the effect of age on condition appears small on an annual basis, the cumulative impact over several decades can significantly compromise a bridge's integrity.
+This underscores the importance of regular inspections and maintenance programs to mitigate aging's effects on bridge conditions.
 
@@
 ## Age Distribution Patterns {#sec-age-distribution}
-The distribution of bridge ages at inspection shows distinct peaks approximately every 25 years, indicating a recurring cycle of bridge construction and major renovation.
+@fig-age-histogram shows the distribution of bridge ages at inspection has distinct peaks approximately every 25 years, indicating a recurring cycle of bridge construction and major renovation.
 This pattern suggests that every 25 years, a significant number of bridges reach the point where extensive maintenance or replacement is necessary.
 The previous peak occurred roughly 25 years ago, meaning the next wave of large-scale repairs or new construction is approaching.
-Recognizing this pattern highlights the importance of early planning and budgeting to address the anticipated surge in infrastructure needs effectively, ensuring resources are ready to manage the simultaneous demands of aging bridges.
+Recognizing this pattern emphasizes the need for proactive planning and budgeting to address the anticipated surge in infrastructure needs, ensuring resources are available to manage the simultaneous demands of aging bridges.
 

### 2024-12-04 other/datasheet/datasheet.qmd
Final prose update

@@
 The "New York Bridges 2016" dataset was created to document and assess the structural conditions and administrative details of bridges across New York State, supporting infrastructure management, maintenance prioritization, and policy-making.
-Developed by the New York State Department of Transportation (NYSDOT) and made publicly available through the Data and Story Library (DASL), this dataset addresses the need for standardized, comprehensive data on bridge deterioration influenced by geographic, administrative, and traffic-related factors.
-Funded as part of New York State’s infrastructure monitoring programs, the dataset serves as a vital resource for researchers, policymakers, and engineers, enabling data-driven decision-making to enhance public safety and infrastructure sustainability.
+Developed by the New York State Department of Transportation (NYSDOT) and made publicly available through the Data and Story Library (DASL), this dataset addresses the need for data on bridge deterioration influenced by geographic, administrative, and traffic-related factors.
+Funded as part of New York State’s infrastructure monitoring programs, the dataset serves as a vital resource for researchers, policymakers, and engineers, enabling informed decision-making to enhance public safety and infrastructure sustainability.
 
@@
 
-The dataset is comprehensive rather than a sample, aiming to capture all operational bridges in New York State as of 2016.
-This ensures extensive geographic and administrative coverage, from urban centers with high traffic density to rural bridges with lower traffic loads.
+The dataset aims to include all operational bridges in New York State as of 2016, providing broad geographic and administrative coverage, from urban centers with high traffic density to rural bridges with lower traffic loads.
 While the data is representative, rural bridges may have less frequent or less detailed inspections, potentially introducing a minor bias.
@@
 The dataset titled "New_York_Bridges_2016" was originally compiled by the New York State Open Data platform[@ny_open_data_2020], with county-level contributions from the New York State Department of Transportation [@county_data].
-Accessed via the Data And Story Library [@dasl_data_2016], which archives historical datasets for academic and public use, it represents a comprehensive resource for understanding bridge conditions across New York State.
+Accessed via the Data And Story Library [@dasl_data_2016], which archives historical datasets for academic and public use, it represents a resource for understanding bridge conditions across New York State.
 
@@
 NYSDOT ensured the reliability of these processes through rigorous training programs for inspectors and periodic calibration of measurement tools.
-The dataset is comprehensive, capturing all operational bridges in the state as of 2016, rather than representing a sample, and thus no specific sampling strategy was required.
+The dataset capturing all operational bridges in the state as of 2016, rather than representing a sample, and thus no specific sampling strategy was required.
 
@@
 In summary, the New York Bridges 2016 dataset was compiled through standardized field inspections and administrative records, validated by trained professionals, and ethically sound due to its exclusive focus on public infrastructure.
-Its methodology ensures high-quality, accurate data for research, policy planning, and infrastructure management, while DASL plays a critical role in archiving and maintaining this resource for public and academic use.
+Its methodology ensures high-quality, accurate data for research, policy planning, and infrastructure management, while DASL plays a role in archiving and maintaining this resource for public and academic use.
 
@@
 The dataset helps identify structurally deficient bridges, prioritize infrastructure projects, and allocate federal funds, such as the $2 billion bridge formula funds under the Infrastructure Investment and Jobs Act (IIJA).
-This dataset is well-suited for tasks like policy development, maintenance scheduling, risk assessment, and infrastructure planning, offering actionable insights into bridge management.
+This dataset is well-suited for tasks like policy development, maintenance scheduling, risk assessment, and infrastructure planning, offering actionable information into bridge management.
 However, its composition reflects current bridge conditions and may not fully account for recently completed repairs, which could influence its applicability for real-time analyses.
 The dataset is tailored to infrastructure evaluations and is not recommended for unrelated uses.
-Overall, it serves as a critical resource for understanding and improving bridge conditions in New York State.
+Overall, it serves as a resource for understanding and improving bridge conditions in New York State.
 

### 2024-12-04 paper/paper.qmd
Final prose update

@@
 ## Overview
-The objective of this methodology is to design a thorough system for collecting and analyzing traffic flow data across New York State bridges.
+The objective of this methodology is to design a system for collecting and analyzing traffic flow data across New York State bridges.
 This additional data addresses an essential limitation in the current dataset, which does not include traffic intensity—a key factor influencing bridge deterioration.
@@
 
-By incorporating these dimensions, the target population captures the variability in bridge conditions and usage, enabling a detailed analysis of how traffic impacts structural integrity.
+By incorporating these dimensions, the target population captures the variability in bridge conditions and usage, enabling an analysis of how traffic impacts structural integrity.
 
